\Title

### 2021-03-19 latex/paper/main.tex
Added VNHC description to Acrobot section

While the simplest model of a gymnast is the variable-length pendulum, a
more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
Here, the top link represents the torso, while the bottom link represents
the legs. 
The acrobot is actuated exclusively at the center joint (the hips)
Controlling the acrobot is a nontrivial task because it is not feedback
linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
To solve the swingup problem, one might begin by designing a leg controller
which provably injects energy into the acrobot, so that the resulting motion
mimics that of a human performing a giant.

@@

Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and they have shown marked improvements to the field of bipedal locomotion 
\cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Indeed, they produce more robust bipedal motion than
controllers which do not depend on velocity \cite{nhvc_incline_walking}.
In particular, VNHCs may be capable of injecting and
dissipating energy from a system in a robust manner, all while producing
realistic biological motion.

@@

The notion of a nonholonomic constraint which can be stabilized via state feedback
was first described by Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
Horn et. al later extended their results in
\cite{hybrid_zero_dynamics_bipedal_nhvcs} and
derived the dynamics of constrained systems.
In this section we reformulate these ideas for the Hamiltonian framework,
because the theory is cleaner when using unactuated and actuated momenta.
For this reason, we take the system of inquiry to be a Hamiltonian
mechanical system in simply actuated coordinates, as in
\eqref{eqn:simple-hamiltonian}.
For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\),
where we now have \((q,p) = (q_u,q_a,p_u,p_a) \in \mathcal{Q}\times \R^n\).

@@

\begin{defn}\label{defn:energy-gain}
    Let \(\mathcal{Q}\) be an
    \(n\)-dimensional smooth manifold.
    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
    field and let \(D \subset M\) be open.
    The system described by \(\dot{x} = f(x)\) 
    \textit{gains energy on \(D\)} if, 
    for all compact sets \(K \subset D\) and for almost every initial
    condition \(x(0) \in K\), there exists \(T > 0\) such
    that \(x(t) \notin K \, (\forall t > T)\).
    The system \textit{loses energy on \(D\)} if it gains energy in
    negative-time.
\end{defn}

@@

%/========== Acrobot ==========/%
\section{The Acrobot VNHC}\label{sec:acrobot}
%specialize the theory of VNHCs to the acrobot. consider constraints which depend
%only on pu, and show arctan vnhc and reduced dynamics. 
%give definition of energy injection/dissipation. conclude with our
%theorem. Make this section short and sweet.
%
%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
%injection for giant stabilization)

For our model of the acrobot, we assume the acrobot is comprised of two massless
rods of equal length \(l\), with equal point masses \(m\) at the tips, as in Figure
\ref{fig:simple-acrobot-model}.
Note that a real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor; 
for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
\(Q_a \in [\frac{\pi}{2}, \pi[\). 
We also ignore any frictional forces at both the hip and shoulder joints.

@@

This system has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
~\eqref{eqn:acrobot-hamiltonian}, where
for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
:= \Sone \times \Sone\), and
\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
:= \R \times \R\).
Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
energy-injecting controller.
Since we need our VNHC to be regular, the following proposition will prove
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
    with \(f \in C^2\left(\mathcal{P}_u; \mathcal{Q}_a\right)\) is a regular
    VNHC of order 1 for the simple acrobot.
\end{prop}
\begin{proof}
    The regularity matrix for this VNHC evaluates to
    \(\frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}\).
    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
\end{proof}

>>>>

While the simplest model of a gymnast is the variable-length pendulum, a
more realistic model is the two-link acrobot (Figure \ref{fig:acrobot}).
Here, the top link represents the torso, while the bottom link represents
the legs. 
The acrobot is actuated exclusively at the center joint (the hips).
Controlling the acrobot is a nontrivial task because it is not feedback
linearizable \cite{nonlinear_controllers_nonintegrable_acrobot}. 
To solve the swingup problem, one might begin by designing a leg controller
which provably injects energy into the acrobot, so that the resulting motion
mimics that of a human performing a giant.

@@

Virtual nonholonomic constraints (VNHCs) have been used for human-robot interaction
\cite{vnhc_human_robot_cooperation,psd_based_vnhc_redundant_manipulator,haptic_vnhc},
error-reduction on time-delayed systems \cite{vnhc_time_delay_teleop},
and they have shown marked improvements to the field of bipedal locomotion 
\cite{nhvc_dynamic_walking,
hybrid_zero_dynamics_bipedal_nhvcs,output_nhvc_bipedal_control}.
Indeed, they produce more robust walking motion in biped robots than
other controllers which do not depend on velocity \cite{nhvc_incline_walking}.
In particular, VNHCs may be capable of injecting and
dissipating energy from a system in a robust manner, all while producing
realistic biological motion.

@@

The notion of a nonholonomic constraint which can be stabilized via state feedback
was first described by Griffin and Grizzle in \cite{nhvc_dynamic_walking}.
Horn et. al later extended their results in
\cite{hybrid_zero_dynamics_bipedal_nhvcs} and
derived the dynamics of constrained systems.
In this section we reformulate these ideas for the Hamiltonian framework,
because the theory is cleaner when using unactuated and actuated momenta.
For this reason, we take the system of inquiry to be a Hamiltonian
mechanical system in simply actuated coordinates, as in
\eqref{eqn:simple-hamiltonian}.
For simplicity of notation, we relabel \((\tilde{q},\tilde{p})\) as \((q,p)\).

@@

\begin{defn}\label{defn:energy-gain}
    Let \(\mathcal{Q}\) be an
    \(n\)-dimensional smooth generalized cylinder.
    Let \(f : \mathcal{Q} \rightarrow \mathcal{Q}\times\R^n\) be a smooth vector
    field and let \(D \subset M\) be open.
    The system described by \(\dot{x} = f(x)\) 
    \textit{gains energy on \(D\)} if, 
    for all compact sets \(K \subset D\) and for almost every initial
    condition \(x(0) \in K\), there exists \(T > 0\) such
    that \(x(t) \notin K \, (\forall t > T)\).
    The system \textit{loses energy on \(D\)} if it gains energy in
    negative-time.
\end{defn}

@@

%/========== Acrobot ==========/%
\section{The Acrobot VNHC}\label{sec:acrobot}
%specialize the theory of VNHCs to the acrobot. consider constraints which depend
%only on pu, and show arctan vnhc and reduced dynamics. 
%give definition of energy injection/dissipation. conclude with our
%theorem. Make this section short and sweet.
%
%TODO: after writing S2, re-introduce what we want to do with acrobot (energy
%injection for giant stabilization)
Recall that we wish to design a VNHC which injects energy into the acrobot
through giant-like motion.
For our model of the acrobot, we assume the torso and leg rods are of equal
length \(l\) with equal point masses \(m\) at the tips, as in Figure
\ref{fig:simple-acrobot-model}.
Note that a real gymnast cannot swing their legs in full circles, though they
are usually flexible enough to raise them parallel to the floor; 
for this reason, we assume that \(q_a \in [-Q_a, Q_a]\) where 
\(Q_a \in [\frac{\pi}{2}, \pi[\). 
We also ignore any frictional forces at both the hip and shoulder joints.

@@

This system has inertia matrix 
\(M\), potential function \(V\) (with respect to
the horizontal bar), and input matrix \(B\) given as follows:
\begin{align}\label{eqn:acrobot-inertia}
    M(q) &= \begin{bmatrix}
        ml^2\left(3+2\cos(q_a)\right) & 
        ml^2\left(1+\cos(q_a)\right) \\
        ml^2\left(1+\cos(q_a)\right) &
        ml^2
    \end{bmatrix} 
    , \\
    \label{eqn:acrobot-potential}
    V(q) &= -mgl\left(2\cos(q_u)+\cos(q_u+q_a)\right)
    , \\
    \label{eqn:acrobot-B}
    B &= [0;1]
    .
\end{align}
The conjugate of momenta is \(p = (p_u,p_a) = M(q)\dot{q}\).
The dynamics of the acrobot in \((q,p)\) coordinates are given in
~\eqref{eqn:acrobot-hamiltonian}, where
for shorthand, we write \(c_u := \cos(q_u)\), \(c_a := \cos(q_a)\), and 
\(c_{ua} := \cos(q_u + q_a)\); likewise, \(s_u := \sin(q_u)\), 
\(s_a := \sin(q_a)\), and \(s_{ua} := \sin(q_u + q_a)\).
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \Sone \times \Sone\), and
\(\mathcal{P} = \R \times \R\).
Hence, we can apply the theory of VNHCs from Section \ref{sec:vnhc} to design an
energy-injecting controller.
Since we need our VNHC to be regular, the following proposition will prove
useful.
\begin{prop}\label{prop:acrobot-fpu-regular}
    A relation \(h(q,p) = q_a - f(p_u) = 0\) 
    with \(f \in C^2\left(\R; \Sone\right)\) is a regular
    VNHC of order 1 for the simple acrobot.
\end{prop}
\begin{proof}
    The regularity matrix for this VNHC evaluates to
    \(\frac{(1+c_a)\partial_{q_u}f(q_u,p_u) + (3+2c_a)}{ml^2(2-c_a^2)}\).
    Since \(\partial_{q_u} f = 0\), this matrix is strictly positive for all values
    of \(q_a\), and hence is full rank everywhere on the constraint manifold.
\end{proof}

To design our VNHC, we begin by examining a person on a
seated swing.
The person extends their legs when the swing moves forwards, and retracts them
when the swing moves backwards.
As the swing gains speed, the person leans their body back while
extending their legs higher, thereby shortening the distance
from their center of mass to the pivot and adding more energy to the swing
\cite{how_to_pump_a_swing}.

Now imagine the person's torso is affixed to the swing's rope so they are
always upright. 
Imagine further that the swing has no seat at all, allowing the person to extend
their legs beneath them. 
This position is identical to that of a gymnast on a bar, which is why we can
use leg motion from the seated swing to design a controller for the acrobot.

The acrobot's legs are rigid rods which cannot retract, so we emulate the person
on a swing by pivoting the legs toward the direction of motion. 
To account for how a person leans back at higher speeds, the legs should pivot to an
angle proportional to the swing's speed.
Since the direction of motion is entirely determined by \(p_u\), 
one such VNHC which emulates this process is \(q_a = \bar{q}_a\arctan( I p_u)\),
displayed in Figure \ref{fig:qa-arctan}.
Here, \(\bar{q}_a \in ]0,\frac{2 Q_a}{\pi}]\) and \(I \in \R\) is a fixed
control parameter.

This constraint does not perfectly recreate giant motion, during which
the gymnast's legs are almost completely extended \cite{usagym_giant}
-- it instead pivots the legs partially during rotations.
However, the behaviour looks similar enough that the constraint should provide a
reasonable foundation for injecting energy into the acrobot.
Hence, our VNHC is
\begin{equation}\label{eqn:acrobot-constraint}
    h(q,p) = q_a - \bar{q}_a \arctan(I p_u)
    .
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{qa_arctan.pdf}
    \caption{The acrobot constraint \(q_a = \bar{q}_a \arctan(I p_u)\).}
    \label{fig:qa-arctan}
\end{figure}

It is not hard to show that the constrained dynamics for the acrobot under
~\eqref{eqn:acrobot-constraint} are
\begin{equation}\label{eqn:acrobot-constrained-dynamics}
    \begin{cases}
    \dot{q}_u &= \frac{(1+I^2 p_u^2)p_u + m^2gl^3\bar{q}_a I(2s_u + s_{ua})(1+c_a) }
            {ml^2(1+I^2 p_u^2)(3+2c_a)}
    ,    \\
    \dot{p}_u &= - m g l (2s_u + s_{ua})
    ,
    \end{cases}
\end{equation}
subject to \(q_a = \bar{q}_a \arctan(I p_u)\).
